Stability in human interaction networks: sector relative sizes, prominence of topological measures and time activity statistics.

### 2015-05-04 paper.tex
contemplando apontamentos do chu

\begin{table}
	\caption{The rescaled circular mean $\theta_\mu'$, the standard deviation $S(z)$, the variance $Var(z)$ and the circular dispersion $\delta(z)$ are given for 1000 uniform distribution simulations within the same number of bins and with the same number of samples. Section~\ref{sec:mtime} describes the theoretical background of directional (or circular) statistics. This typical table was made using LAD list messages, and the results are about the same for other lists, as exposed in Supporting Information. Most homogeneous activity was found in seconds and minutes, where the mean is has no real meaning. Hours of the day exhibited the most concentrated activity (lowest $S(z)$, $Var(z)$ and $\delta(z)$), with mean between 14h and 15h ($\theta'=-9.61$). Weekdays, months and month days have mean very near zero, and considerable dispersion.}
	\begin{center}
    \begin{tabular}{ |l|| c|c|c|c|c||c|c| }
        \hline
%scale & $\theta_\mu'$ & $S(z)$ & $Var(z)$ & $\delta(z)$ & $\frac{max(incidence)}{min(incidence)}$ & $ \mu_{\frac{max(incidence')}{min(incidence')}} $ & $ \sigma_{\frac{max(incidence')}{min(incidence')} } $ \\ \hline\hline
& $\theta_\mu'$ & $S(z)$ & $Var(z)$ & $\delta(z)$  \\ \hline\hline
	\input{tables/tab2TimeLAD__}
    \end{tabular}
\end{center}
\label{tab:circ}
\end{table}

@@

There is a concentration of hub activity and of vertex with few connections. Table~\ref{autores} shows this expected distribution of activity among participants in a scale-free context.

@@

The distribution of vertices in the hubs, intermediary, periphery Erd\"os sectors defined in Section~\ref{sectioning} is remarkably stable along time, provided that a sufficiently large sample of 200 or more messages is considered. 
Moreover, the same distribution applies to the networks of all the four email lists, as demonstrated in the various figures in the Support Information. 
Typically, $\approx [3-12]\%$ of the vertices are found to be hubs, $\approx [15-44]\%$ are intermediary and $\approx [44-82]\%$ are peripheral, which is consistent with the literature~\cite{secFree}.
These results hold for the total, in and out degrees and strengths.
Stable distributions can also be obtained for 100 or less messages if classification of the three sectors is performed with one of the compound criteria established in Section~\ref{sectioning}. 
The networks hold their basic structure with as few as 10-50 messages; concentration of activity and the abundance of low-activity participants take place even with very few messages, which is highlighted in the last figures of the Support Information.
A minimum window size for observation of more general properties might be inferred by monitoring the giant component and the degeneration of the Erd\"os sectors.

@@

The first principal component is an average of centrality metrics: degrees, strengths and betweenness centrality. Therefore, all of these centrality measurements are equally important for characterizing the networks. On one hand, the relevance of all centrality metrics is not surprising since they may be highly correlated. The degree and strength, for instance, are highly correlated, with Spearman correlation coefficient $\in [0.95,1]$ and Pearson coefficient $\in [0.85,1)$ for $ws>1000$.
On the other hand, each measure relates to a different participation characteristic, and their equal relevance is noticeable.
The clustering coefficient is presented in almost perfect orthogonality to centrality measures.
\begin{figure*} 
   \centering
        \includegraphics[width=.6\textwidth,height=10cm]{figs/im13PCAPLOT_}
	\caption{Prevalence of the symmetry-related measures over clustering for dispersion of participants in the space of topological measures. This is an important result from the PCA analysis, together with stability of component formation, and can be observed in Table~\ref{tab:pcain}.
		The first plot shows degree versus clustering coefficient.
		The second plot is similar, but the first component is an average of centrality metrics.
		The third plot resulted from applying PCA with symmetry-related metrics, where greater dispersion in the second component can be observed among all sectors.
		In this case, the second component is representative of symmetry measurements of vertex interactions and clustering coefficient is only relevant for the third component.
		Greater dispersion suggests symmetry-related metrics as being more powerful for characterizing interaction networks than clustering coefficient, especially for hubs and intermediary vertices.
		This figure was done with a snapshot of the LAU list in a window size of $ws = 1000$ messages.
		The general layout is recurrent and consistent with the literature: most connected vertices have low clusterization while higher clusterization is gradually more incident as the number of connections is lowered.
		Similar structures were observed in all window sizes $ws\;\in\;[500,10000]$ and for networks of other email lists, which points to a common relationship between the measures of degree, strength and betweenness centrality, the measures of symmetry and the measure of clustering coefficient described in Section~\ref{measures}.
    }
    \label{fig:sym}
\end{figure*}

@@

Let a sector to which a vertex belongs yield a type for the corresponding participant. Accordingly, a participant may be considered peripheral, intermediary or a hub. This trivial consideration raises important questions: is this typology stigmatizing or prejudice inclined in any way? What do we know about each type and about the typology itself?
These general considerations about this primitive typology will help us answer the first questions:
\begin{itemize}
	\item A participant belongs to many networks (his family, family of a friend, the email list to which he belongs, work friends, etc.).
	\item A participant might belong to all three sectors at the same time. Actually, it is reasonable to assume that almost every person belongs to all three sectors for some of the networks.
	\item The participant often transitions from one sector to another within a network.
	\item Reported stability of network structure is concomitant with the continuous change of participants and the type of each participant.
\end{itemize}

@@

This typology, derived entirely from objective metrics, is a bridge from exact to human sciences. Obviously, the analysis may be enriched by using concepts from other typologies, such as Meyer-Briggs, Pavlov or the authoritarian types of the F-Scale~\cite{adorno}.

>>>>

\begin{table}
	\caption{The rescaled circular mean $\theta_\mu'$ and the circular dispersion $\delta(z)$ described in Section~\ref{sec:mtime}. This typical table was made using all LAD list messages, and the results are the same for other lists, as exposed in Supporting Information. Most uniform distribution of activity was found in seconds and minutes, where the mean has little meaning. Hours of the day exhibited the most concentrated activity (lowest $\delta(z)$), with mean between 14h and 15h ($\theta'=-9.61$). Weekdays, month days and months have mean near zero (i.e. near the beginning of the week, month and year) and high dispersion.}
	\begin{center}
    \begin{tabular}{ |l|| c|c| }
        \hline
%scale & $\theta_\mu'$ & $S(z)$ & $Var(z)$ & $\delta(z)$ & $\frac{max(incidence)}{min(incidence)}$ & $ \mu_{\frac{max(incidence')}{min(incidence')}} $ & $ \sigma_{\frac{max(incidence')}{min(incidence')} } $ \\ \hline\hline
%& $\theta_\mu'$ & $S(z)$ & $Var(z)$ & $\delta(z)$  \\ \hline\hline
& $\theta_\mu'$ & $\delta(z)$  \\ \hline\hline
	\input{tables/tab2TimeLAD___}
    \end{tabular}
\end{center}
\label{tab:circ}
\end{table}

@@

There is a concentration of hub activity and an abundance of vertex with few connections. Table~\ref{autores} shows this expected distribution of activity among participants in a scale-free context.

@@

The distribution of vertices in the hubs, intermediary, periphery Erd\"os sectors defined in Section~\ref{sectioning} is remarkably stable along time, provided that a sufficiently large sample of 200 or more messages is considered. 
Moreover, the same distribution applies to the networks of all the four email lists, as demonstrated in Figure~\ref{fig:sectIL} and the Support Information. 
Typically, $\approx [3-12]\%$ of the vertices are found to be hubs, $\approx [15-44]\%$ are intermediary and $\approx [44-82]\%$ are peripheral, which is consistent with the literature~\cite{secFree}.
These results hold for the total, in and out degrees and strengths.
Stable distributions can also be obtained for 100 or less messages if classification of the three sectors is performed with one of the compound criteria established in Section~\ref{sectioning}. 
The networks hold their basic structure with as few as 10-50 messages; concentration of activity and the abundance of low-activity participants take place even with very few messages, which is highlighted in the last figures of the Support Information.
A minimum window size for observation of more general properties might be inferred by monitoring the giant component and the degeneration of the Erd\"os sectors.

@@

The first principal component is an average of centrality metrics: degrees, strengths and betweenness centrality. Therefore, all of these centrality measurements are equally important for characterizing the networks. On one hand, the relevance of all centrality metrics is not surprising since they may be highly correlated. The degree and strength, for instance, are highly correlated, with Spearman correlation coefficient $\in [0.95,1]$ and Pearson coefficient $\in [0.85,1)$ for $ws>1000$.
On the other hand, each measure relates to a different participation characteristic, and their equal relevance is noticeable.
The clustering coefficient is presented in almost perfect orthogonality to centrality measures.
\begin{figure*} 
   \centering
        \includegraphics[width=.6\textwidth,height=10cm]{figs/im13PCAPLOT_}
	\caption{The first plot shows degree versus clustering coefficient.
		In the second plots, the first component is an average of centrality metrics, but the second component is still clustering coefficient.
		The third plot exhibits the greater dispersion in the symmetry-related second component.
		In this case, clustering coefficient is only relevant for the third component.
		Greater dispersion suggests that symmetry-related metrics are more powerful for characterizing interaction networks than clustering coefficient, especially for hubs and intermediary vertices.
		This general layout is recurrent and consistent with the literature: most connected vertices have low clusterization while higher clusterization is gradually more incident as the number of connections is lowered.
		This prevalence of the symmetry-related measures over clustering coefficient for the dispersion of participants in the space of topological measures is an important result from the PCA analysis, together with temporal stability of component formation, and can be observed in Table~\ref{tab:pcain}.
		This figure was done with a snapshot of the LAU list in a window size of $ws = 1000$ messages.
		Similar structures were observed in all window sizes $ws\;\in\;[500,10000]$ and for networks of other email lists, which points to a common relationship between the measures of degree, strength and betweenness centrality, the measures of symmetry and the measure of clustering coefficient described in Section~\ref{measures}.
    }
    \label{fig:sym}
\end{figure*}

@@

Let a sector to which a vertex belongs yield a type for the corresponding participant. Accordingly, a participant may be considered peripheral, intermediary or a hub. This trivial consideration raises important questions: is this typology stigmatizing or prejudice inclined in any way? What else do we know about each type and about the typology itself?
These general considerations about this primitive typology will help us answer the first questions:
\begin{itemize}
	\item A participant belongs to many networks (his family, family of a friend, the email list to which he belongs, work friends, etc.).
	\item A participant might belong to all three sectors at the same time. Actually, it is reasonable to assume that almost every person belongs to all three sectors for some of the networks.
	\item The participant often transitions from one sector to another within a network.
	\item Reported stability of network structure is concomitant with the continuous change of participants and the type of each participant.
\end{itemize}

@@

This typology is a bridge from exact to human sciences and may be enriched with concepts from other typologies, such as Meyer-Briggs, Pavlov or the authoritarian types of the F-Scale~\cite{adorno}.
